Calibrating the pixel-level Kepler imaging data with a causal data-driven model \\

### 2015-02-25 cpm.tex
up to page 10

The second general idea about data-driven modeling is the investigator's beliefs
  about the causal processes that generate the data
  are crucial in restricting the kinds of data that can be used as input to the model.
That is, different assumptions about the physical properties of the data
  and the data-taking system
  lead to different structures for the data-driven model.
In the case of \Kepler\ data, if we examine two arbitrary stars far away from each other 
on the same CCD as in Fig~\ref{ccd} (every grid in the plot is a pixel light curve time 
series), we can obviously see that the two stars that are hundreds pixel away from each 
other do have similar trends. If we believe that different stars in the \Kepler\ field vary 
independently (that is, are not physically synchronized in any way) because of the 
distance between them, then the only reason that one star might show variability that is strongly \emph{predictive} (useful for prediction) of another star's variability
  is that both stars are being observed by the same device or spacecraft.
That is, one star's pixels can be used to predict another star's pixels
  inasmuch as spacecraft issues imprint on both stars in related ways;
  they share a common cause---the systematics.  
For another example, if we think the spacecraft is being affected by
  processes that take place over time-scales longer than a single read-out
  (for example, thermal processes),
  then it would be sensible to model the data at time $t$ using not just simultaneous data, but also data from a range of times around $t$.
For another example, if an investigator doesn't care about preserving stellar variability,
  and just wants to detect exoplanet transits (say),
  pixels from the target star can be used to predict pixels from the target star,
  provided they are at large-enough time lags that they don't overlap (in time)
  the signals of interest.
That is, if the model is designed to fit not just spacecraft variability
  but also intrinsic stellar variability,
  the predictive model will be permitted to use as input pixels that \emph{do} overlap the 
  target star.
In what follows, we are going to use input data both from pixels of other stars and the  
target star pixels' past and furture.
These models ought to remove both spacecraft variability and intrinsic stellar variability, 
which is optimal for exoplanet transit searching.

@@

The fourth general idea involved in this kind of data-driven modeling is that the models aren't \emph{interpretable}.
This means, in particular, that although the model might do a good job of \emph{predicting} the target pixels,
  using a linear (or more complex) combination of the input pixel data,
  it won't deliver anything that can be unambiguously interpreted as the \emph{flux} of the star in question,
  or any other signal we care about.
The data-driven model \emph{effectively} describes the pointing, point-spread function, and flat-field
  of the telescope and camera,
  plus the variability of every star and every exoplanet transit,
  but it does so without ever \emph{explicitly} creating any of those objects.
We have to make some kind of heuristic or interpretive move to extract from the data-driven model the quanitites of interest.

@@

@@

Let's consider a particular pixel $m$ in the focal plane in one particular month
(here we always use data per month rather than quarter,  
since there is discontinuity between every 30-ish-day period).
In that month, the spacecraft delivers $N\sim 1300$ measurements $I_{mn}$
  of the intensity falling in that pixel at the $N$ times $t_n$ at which exposures were taken during the month.
(For intensity measurements $I_{mn}$ here we are using pixels from the Target Pixel Files.)
We want to build a prediction for the intensity $I_{mn}$ in pixel $m$ at each time $t_n$
  using the intensities $I_{m'n}$ of other pixels $m'$.
The question is:  What other pixels to choose?
There are many possible qualitatively and quantitatively different choices here.
Somewhat arbitrarily,
  we choose all the pixels $m'$ associated with the $Q$ stars on the same CCD in the same quarter
  that are closest in magnitude
  (\Kepler\ magnitude as reported in the \Kepler\ Input Catalog)
  to the target star associated with pixel $m$.
This set of pixels%
  ---from the $Q$ stars on the same CCD---%
  is the \emph{pixel set} $\set{M}_m$ associated with pixel $m$.
   To rule out any direct optical cross-talk by starry 
light, we require that the predictor pixels are from stars
sufficiently far away from the target star (at least 20 pixels distance on the CCD).
Note that because the set $\set{M}_m$ is of pixels associated with \emph{different} stars
  than the star associated with pixel $m$, and are far from it on CCD, 
  the pixel $m$ will not be in the set $\set{M}_m$,
  and nor will any of pixel $m$'s close neighbors.
That is, there will be no (or almost no) overlap in stellar illumination of pixel $m$
  and the pixels $m'\in\set{M}_m$.

@@

In addition to pixel set $m'\in\set{M}_m$ from other stars,  
  we also include past and future of the target star --- autoregressive (AR) component as input 
  to remove more of the stellar variability and thus increase the sensitivity for transits. 
To do this, we select an exclusion window of half-width $\Delta t$ including R data points, 
  where $\Delta t\sim 10$ (as big as the window in train-and-test framework) 
  around the point of time $t_{n}$ being corrected, 
  to ensure that we do not remove the transit itself and then use S closest future and past time points. 
That is, for every pixel m in the target star , we construct S virtual time series, 
  in which every component v is defined to be     
\begin{eqnarray}
I_{vn} = I_{mn-R-k}\ or\ I_{vn} = I_{mn+R+k},
\quad,
\end{eqnarray}
where k = 0, 1, ..., S, so if there are totally M pixels in the target stars, 
  we will finally add $2\cdot M\cdot S$ autoregressive components into the predictors, 
  which constructs the autoregressive set $\set{V}_m$

As mentioned above, in the context of this project,
  a model is something that can predict data given settings of some parameters,
  and an objective function that can be used to find best values for those parameters.
We prefer to work in a probabilistic mode in which the objective function can be justified
  in terms of a likelihood (probability for the data given the model)
  or a posterior probability density function (likelihood times a prior pdf).
The \name\ treats each \Kepler\ data point $I_{mn}$ as being
  predictable from a linear combination of data points $I_{m'n}$,
  where $m'$ is from the set of non-overlapping pixels $m'\in\set{M}_m$.
\begin{eqnarray}
I_{mn}         &=& I^{\ast}_{mn} + e_{mn}
\\
I^{\ast}_{mn}  &=& \sum_{m'\in\set{M}_m} a_{mnm'}\,I_{m'n} + \sum_{v\in\set{V}_m} b_{mnv}\,I_{vn}
\quad,
\end{eqnarray}
where $I^{\ast}_{mn}$ is the prediction (by the model) for data point $I_{mn}$,
  $e_{mn}$ is a noise contribution or residual away from the prediction,
  and the $a_{mnm'}$ and $b_{mnk}$ are the parameters (linear coefficients of the prediction).
The parameters $a_{mnm'}$ and $b_{mnk}$ have indices $m$ because
  they are different for every pixel $m$,
  and they will even be different for every time step $n$;
  that is, there will be a separate best-fit value for the parameters for every $I_{mn}$.
If we presume that the residuals away from the prediction are normally distributed with zero mean
  and known variance,
  likelihood optimization reduces to chi-squared minimization.
We add to the standard chi-squared definition a regularization term
  (equivalent to multiplying the likelihood by a Gaussian prior pdf)
  that penalizes large absolute values for the the coefficients $a_{mnm'}$:
\begin{eqnarray}
\chi^2_{mn}    &=& \sum_{n'\in\set{N}_n} \frac{[I_{mn'} - I^{\ast}_{mnn'}]^2}{\sigma^2_{mn'}}
                 + \lambda_{a}\sum_{m'\in\set{M}_m}a_{mnm'}^2 + \lambda_{b}\sum_{v\in\set{V}_m}b_{mnv}^2
\\
I^{\ast}_{mnn'} &=& \sum_{m'\in\set{M}_m} a_{mnm'}\,I_{m'n'} + \sum_{v\in\set{V}_m} b_{mnv}\,I_{vn'}
\quad,
\end{eqnarray}
where the $\sigma^2_{mn'}$ are the (presumed known) individual-pixel noise variances,
  and $\lambda_{a}$, $\lambda_{b}$ set the strength of the regularization (or width of the prior pdf) for parameters $a_{mnm'}, b_{mnv}$.

>>>>

The second general idea about data-driven modeling is the investigator's beliefs
  about the causal processes that generate the data
  are crucial in restricting the kinds of data that can be used as input to the model.
That is, different assumptions about the physical properties of the data
  and the data-taking system
  lead to different structures for the data-driven model.
In the case of \Kepler\ data, if we examine two arbitrary stars far away from each other 
on the same CCD as in Fig~\ref{ccd} (every grid in the plot is a pixel light curve time 
series), we can obviously see that the two stars that are hundreds pixel away from each 
other do have similar trends. If we believe that different stars in the \Kepler\ field vary 
independently (that is, are not physically synchronized in any way) because of the 
distance between them, then the only reason that one star might show variability that is strongly \emph{predictive} (useful for prediction) of another star's variability
  is that both stars are being observed by the same device or spacecraft.
That is, one star's pixels can be used to predict another star's pixels
  inasmuch as spacecraft issues imprint on both stars in related ways;
  they share a common cause---the systematics.  
For another example, if we think the spacecraft is being affected by
  processes that take place over time-scales longer than a single read-out
  (for example, thermal processes),
  then it would be sensible to model the data at time $t$ using not just simultaneous data, but also data from a range of times around $t$.
For another example, if an investigator doesn't care about preserving stellar variability,
  and just wants to detect exoplanet transits (say),
  pixels from the target star can be used to predict pixels from the target star,
  provided they are at large-enough time lags that they don't contain information about
  the signals of interest (i.e., the transits).
That is, if the model is designed to fit not just spacecraft variability
  but also intrinsic stellar variability,
  the predictive model will be permitted to use as input pixels that \emph{do} overlap the 
  target star.
In what follows, we are going to use input data both from pixels of other stars and the  
target star pixels' past and furture.
These models ought to remove both spacecraft variability and intrinsic stellar variability, 
which is optimal for exoplanet transit searching.

@@

The fourth general idea involved in this kind of data-driven modeling is that the models often aren't \emph{interpretable}, or at least hard to interpret.
This means, in particular, that although the model might do a good job of \emph{predicting} the target pixels,
  using a linear (or more complex) combination of the input pixel data,
  it won't deliver anything that can be unambiguously interpreted as the \emph{flux} of the star in question,
  or any other signal we care about.
The data-driven model \emph{effectively} describes the pointing, point-spread function, and flat-field
  of the telescope and camera,
  plus the variability of every star and every exoplanet transit,
  but it does so without ever \emph{explicitly} creating any of those objects.
We have to make some kind of heuristic or interpretive move to extract from the data-driven model the quanitites of interest.

@@

\Bernhard{Hogg: maybe you want to add sth about the device being 'self-calibratable'?}

@@

Let's consider a particular pixel $m$ in the focal plane in one particular month
(here we always use data per month rather than quarter,  
since there is discontinuity between every 30-ish-day period).\Bernhard{as an aside: I wonder whether \name reduces these discontinuities at all? (if we were to apply it across the discontinuities)}
In that month, the spacecraft delivers $N\sim 1300$ measurements $I_{mn}$
  of the intensity falling in that pixel at the $N$ times $t_n$ at which exposures were taken during the month.
(For intensity measurements $I_{mn}$ here we are using pixels from the Target Pixel Files.\Bernhard{include reference})
We want to build a prediction for the intensity $I_{mn}$ in pixel $m$ at each time $t_n$
  using the intensities $I_{m'n}$ of other pixels $m'$.
The question is:  What other pixels to choose?
There are many possible qualitatively and quantitatively different choices here.
Assuming that they are affected by similar systematic errors as the target star pixels, we choose all the pixels $m'$ associated with the $Q$ stars on the same CCD in the same quarter
  that are closest in magnitude
  (\Kepler\ magnitude as reported in the \Kepler\ Input Catalog)
  to the target star associated with pixel $m$.
This set of pixels%
  ---from the $Q$ stars on the same CCD---%
  is the \emph{pixel set} $\set{M}_m$ associated with pixel $m$.
   To rule out any direct optical cross-talk by starry 
light, we require that the predictor pixels are from stars
sufficiently far away from the target star (at least 20 pixels distance on the CCD).
Note that because the set $\set{M}_m$ is of pixels associated with \emph{different} stars
  than the star associated with pixel $m$, and are far from it on CCD, 
  the pixel $m$ will not be in the set $\set{M}_m$,
  and nor will any of pixel $m$'s close neighbors.
That is, there will be no (or almost no) overlap in stellar illumination of pixel $m$
  and the pixels $m'\in\set{M}_m$.

@@

In addition to pixel set $m'\in\set{M}_m$ from other stars,  
  we also include past and future of the target star --- autoregressive (AR) component as input 
  to remove more of the stellar variability and thus increase the sensitivity for transits. 
To do this, we select an exclusion window of half-width $\Delta t$ including R data points, 
  where $\Delta t\sim 10$\, h (as big as the window in the train-and-test framework) 
  around the point of time $t_{n}$ being corrected, 
  to ensure that we do not remove the transit itself and then use $S$ closest future and past time points. 
That is, for every pixel $m$ in the target star , we construct $S$ virtual time series, 
  in which every component $v$ is defined to be     
\begin{eqnarray}
I_{vn} = I_{mn-R-k}\ or\ I_{vn} = I_{mn+R+k},
\quad,
\end{eqnarray}
\Bernhard{I find the notation $I_{mn-R-k}$ confusing --- better write $I_{m,n-R-k}$, and then for consistency also $I_{v,n}$ etc.}
where $k = 0, 1,\dots, S,$ so if there are totally $M$ pixels in the target stars, 
  we will finally add $2\cdot M\cdot S$\Bernhard{not quite: if I am right, then there are overall $(2R-1)$ in the exclusion zone, and $2S+2$ are used, for each pixel} autoregressive components into the predictors, 
  which constructs the autoregressive set $\set{V}_m$

As mentioned above, in the context of this project,
  a model is something that can predict data given settings of some parameters,
  and an objective function that can be used to find best values for those parameters.
It is often preferable to work in a probabilistic mode in which the objective function can be justified
  in terms of a likelihood (probability for the data given the model)
  or a posterior probability density function (likelihood times a prior pdf).
The \name\ treats each \Kepler\ data point $I_{mn}$ as being
  predictable from a linear combination of data points $I_{m'n}$,
  where $m'$ is from the set of non-overlapping pixels $m'\in\set{M}_m$.
\begin{eqnarray}
I_{mn}         &=& I^{\ast}_{mn} + e_{mn}
\\
I^{\ast}_{mn}  &=& \sum_{m'\in\set{M}_m} a_{mnm'}\,I_{m'n} + \sum_{v\in\set{V}_m} b_{mnv}\,I_{vn}
\quad,
\end{eqnarray}
where $I^{\ast}_{mn}$ is the prediction (by the model) for data point $I_{mn}$,
  $e_{mn}$ is a noise contribution or residual away from the prediction,
  and the $a_{mnm'}$ and $b_{mnk}$ are the parameters (linear coefficients of the prediction).
The parameters $a_{mnm'}$ and $b_{mnk}$ have indices $m$ because
  they are different for every pixel $m$,
  and they will even be different for every time step $n$;
  that is, there will be a separate best-fit value for the parameters for every $I_{mn}$.
If we presume that the residuals away from the prediction are normally distributed with zero mean
  and known variance,
  likelihood optimization reduces to chi-squared \Bernhard{maybe better $\chi^2$?} minimization.
We add to the standard chi-squared definition a regularization term
  (equivalent to multiplying the likelihood by a Gaussian prior pdf)
  that penalizes large absolute values for the the coefficients $a_{mnm'}$:
\begin{eqnarray}
\chi^2_{mn}    &=& \sum_{n'\in\set{N}_n} \frac{[I_{mn'} - I^{\ast}_{mnn'}]^2}{\sigma^2_{mn'}}
                 + \lambda_{a}\sum_{m'\in\set{M}_m}a_{mnm'}^2 + \lambda_{b}\sum_{v\in\set{V}_m}b_{mnv}^2
\\
I^{\ast}_{mnn'} &=& \sum_{m'\in\set{M}_m} a_{mnm'}\,I_{m'n'} + \sum_{v\in\set{V}_m} b_{mnv}\,I_{vn'}
\quad,
\end{eqnarray}
\Bernhard{(1) why does $I$ have three indices now, and what does that mean? In Eq. (3) there are only two indices. (2) do we need the sigmas at all, or could we absord this into the lambdas?}
where the $\sigma^2_{mn'}$ are the (presumed known)\Bernhard{known, or estimated from data? If not estimated, how do we know them?} individual-pixel noise variances,
  and $\lambda_{a}$, $\lambda_{b}$ set the strength of the regularization (or width of the prior pdf) for parameters $a_{mnm'}, b_{mnv}$. In general, $\lambda_a$ and $\lambda_b$ could depend on $m,n$, however, we do not make use of this freedom.
